---
title: "HPMT 5214 - Assignment 2"
author: "Evan Munro"
date: "2/8/2021"
output: 
  prettydoc::html_pretty:
    theme: cayman
    toc: true
---

```{r echo = FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

# Prompt

The historical data of daily days of service at cardiology department from May 2010
through May 2013 is given. Import the data into RStudio and use package fpp2 to do the
following.

```{r}
library(fpp2)
cardio = readxl::read_excel('../Data/HW2cardioData.xlsx') 
cardio = cardio[order(cardio$Date),] # order with oldest observation in row 1
```

***

### **1)** Create a `ts` object from the data (Note: ts object should be based on daily basis).
```{r fig.align = 'center'}
# extract date from datetime 
row1DOY = as.Date(cardio$Date[1], "%Y-%m-%d")
# turn date into day of year 
row1DOY = as.numeric(strftime(row1DOY, format = "%j"))
cardioTS = ts(cardio[,1], start = c(2010, row1DOY), frequency = 365)
```

```{r include=FALSE, eval=FALSE}
# plotting practice just for fun --- not part of assignment
autoplot(cardioTS) + 
  ggtitle("Patients Seen by Cardiology Department") +
  expand_limits(x = c(2010, 2014)) + 
  xlab("Year") +
  ylab("Days of Service (# of patients seen per day)") +
  geom_text(aes(x = 2010 + 121/365, y = 11, label = as.Date(cardio$Date[1], "%Y-%m-%d")), 
            hjust = "right", vjust = "bottom", fontface = "italic", size = 3.5) +
  geom_text(aes(x = 2013 + 151/365, y = 23, label = as.Date(cardio$Date[1127], "%Y-%m-%d")),
            hjust = "left", vjust = "bottom", fontface = "italic", size = 3.5)
```


***

### **2)** Split the data into training and testing. Put the data from May 1, 2010 to Dec 31, 2012 into training set and the rest of the data into testing set.

```{r}
startDOY = as.numeric(strftime("2010-05-01", format = "%j"))
endDOY = as.numeric(strftime("2012-12-31", format = "%j"))
trainSet = window(cardioTS, start = c(2010, startDOY), end = c(2012, endDOY), 
                     frequency = 365)
testSet = window(cardioTS, start = c(2013, 1), frequency = 365)
```


***

### **3)** Use average, naïve and seasonal naïve methods to forecast the testing set.

```{r}
nrow(testSet)
avgForc = meanf(trainSet, h = 152)
naiveForc = naive(trainSet, h = 152)
snaiveForc = snaive(trainSet, h = 152)
```


***

### **4)** Develop the autoplot for the previous step.

```{r fig.align="center"}
autoplot(cardioTS) + 
  autolayer(avgForc, series = "Mean Method", PI = FALSE) +
  autolayer(naiveForc, series = "Naive Method", PI = FALSE) +
  autolayer(snaiveForc, series = "Seasonal Naive Method", PI = FALSE) +
  ggtitle("Forecasts of daily patients seen by Cardiology Department") +
  xlab("Year") +
  ylab("Number of patients seen per day") + 
  guides(colour = guide_legend(title = "Forecast"))
```

***

### **5)** What are the accuracy measures of the forecasting methods used in Part (3) and which method give the most accurate forecasts?  

```{r}
accuracy(avgForc, testSet)
accuracy(naiveForc, testSet)
accuracy(snaiveForc, testSet)
```
The seasonal naive method provided the most accurate forecast, which is evidenced by the test set MAPE, MASE, MAE, and RMSE being lowest for the seasonal naive forecast. 

***


</br>
</br>
</br>
</br>
















