---
title: "HPMT 5214 - Assignment 3"
author: "Evan Munro"
date: "2/15/2021"
output: 
  prettydoc::html_pretty:
    theme: cayman
    toc: true
---

```{r echo = FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

# Prompt

Data set `books` (available in fpp2) contains the daily sales of paperback and hardcover books at
the same store. The task is to forecast the next four days’ sales for paperback and hardcover
books.

```{r}
library(fpp2)
data(books)
str(books)
```

***

### **a)** Plot the series and discuss the main features of the data.
```{r fig.align='center'}
autoplot(books, facets = TRUE) + 
  ggtitle('Daily book sales at a store') +
  xlab("Day") +
  ylab("Books Sold") +
  expand_limits(y = c(100, 300)) +
  scale_x_continuous(breaks=c(1,8, 15, 22, 29, 30))
```

Both the paperback daily sales and hardcover daily sales appear to have a positive trend of increasing daily sales over time. There does not appear to be any weekly seasonality for either paperback or hardcover daily sales. 

```{r fig.align='center'}
autoplot(books) + 
  ggtitle('Daily book sales at a store') +
  xlab("Day") +
  ylab("Books Sold") +
  expand_limits(y = c(100, 300)) +
  scale_x_continuous(breaks=c(1,7, 14, 21, 28, 30)) +
  coord_fixed(0.07)
```

It looks like hardcover daily sales are higher on average than paperback daily sales.

***

### **b)** Use the `ses()` function to forecast each series, and plot the forecasts.

```{r fig.align='center'}
pprbakSES = ses(books[,1], h = 4)
hrdcvrSES = ses(books[,2], h = 4)
autoplot(pprbakSES) + 
  autolayer(fitted(pprbakSES), series = "Paperback SES") + 
  ggtitle("Forecast of daily paperback sales") +
  xlab("Day") +
  ylab("Books Sold")

autoplot(hrdcvrSES) + 
  autolayer(fitted(hrdcvrSES), series = "Hardcover SES") + 
  ggtitle("Forecast of daily hardcover sales") +
  xlab("Day") +
  ylab("Books Sold")
```


***

### **c)** Compute the RMSE values for the training data in each case.

```{r}
accuracy(pprbakSES)[1,'RMSE']
accuracy(hrdcvrSES)[1,'RMSE']
```


***


</br>
</br>
</br>
</br>
















